---
title: "Final Project Work"
author: "Adam Haertter & Brennan Mulligan"
format: html
---

# Initial Steps

```{r lib}
#| message: false
#| echo: false

library("tidyverse")
library("rvest")
library("magrittr")
library("here")
```

## Scraping Data

See our [Python scraper](scripts/scrape_vgchartz.py). VGChartz renders its data through a JavaScript table, so the data itself is not accessible through the native HTML for R scraping.

## Wrangling Data

See [wrangle_data.R](scripts/wrangle_data.R).

> You shouldn't need to run this line unless the `data/wrangled/` dir is empty.

```{r}
source("scripts/wrangle_data.R")
```

## Loading Data into Environment

Rerun this if you need to reset the master table. The dataframe with all info is called `master`

```{r}
load(file = here("data", "wrangled", "wrangled_data.rda"))
```

# Data Analysis

Most popular genres by publisher.

```{r}
#master %>% group_by(Publisher, Genre) %>% 
#  summarize("sum" = sum(Total.Sales)) %>% 
#  drop_na() %>% arrange(sum) %>%
#  filter(sum > 0.00) 

publishers <- master[, c(3,5,8)]  %>% drop_na() %>%
  group_by(Publisher.Simple) %>% summarize("Sum.Sales" = sum(Total.Sales)) %>% arrange(desc(Sum.Sales))

# It appears our biggest publishers are Nintendo, Electronic Arts, Activision, Ubisoft, and Sony
# Since they are our most successful companies, let's look into what genre is their most successful by sales

publisher_by_genre <- function(publisher) {
  master[, c(3,4,5,8)] %>%
  filter(Console != "Series" & Console != "All" & Publisher.Simple == publisher) %>%
  drop_na() %>%
  group_by(Genre) %>% summarize("Sum.Sales" = sum(Total.Sales)) %>% arrange(desc(Sum.Sales))
}

sales_by_genre_plot <- function(publisher) {
  ggplot(publisher_by_genre(publisher), aes(x = reorder(Genre, -Sum.Sales), y = Sum.Sales, fill = Genre)) + geom_bar(stat = "identity", position = position_dodge()) + theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + labs(x = "Genre", y = "Sales", title = paste(publisher, " Sales By Genre"))
}

pub_by_genre_plot <- function(publisher) {
  ggplot(publisher_by_genre(publisher), aes(x = reorder(Publisher.Simple, -Sum.Sales), y = Sum.Sales, fill = Publisher.Simple)) + geom_bar(stat = "identity", position = position_dodge()) + theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + labs(x = "Genre", y = "Sales", title = paste(publisher, "'s Most Developed Genre"))
}

sales_by_genre_plot("Nintendo")
sales_by_genre_plot("Electronic Arts")
sales_by_genre_plot("Activision")
sales_by_genre_plot("Ubisoft")
sales_by_genre_plot("Sony")
```