---
title: "Final Project Work"
author: "Adam Haertter & Brennan Mulligan"
format: html
---

# Initial Steps

```{r lib}
#| message: false
#| echo: false

library("tidyverse")
library("rvest")
library("magrittr")
library("here")
```

## Scraping Data

See our [Python scraper](scripts/scrape_vgchartz.py). VGChartz renders its data through a JavaScript table, so the data itself is not accessible through the native HTML for R scraping.

## Wrangling Data

See [wrangle_data.R](scripts/wrangle_data.R).

> You shouldn't need to run this line unless the `data/wrangled/` dir is empty.

```{r}
source("scripts/wrangle_data.R")
```

## Loading Data into Environment

Rerun this if you need to reset the master table. The dataframe with all info is called `master`

```{r}
load(file = here("data", "wrangled", "wrangled_data.rda"))
```

# Data Analysis

Most popular genres by publisher.

```{r}
#master %>% group_by(Publisher, Genre) %>% 
#  summarize("sum" = sum(Total.Sales)) %>% 
#  drop_na() %>% arrange(sum) %>%
#  filter(sum > 0.00) 

publishers <- master[, c(2,5,7)] %>% drop_na() %>% group_by(Publisher) %>% summarize("Sum.Sales" = sum(Total.Sales)) %>% arrange(Publisher)
```
